---
title: "PSTAT 122 Final Exam"
author: "Justin Lee"
date: "6/11/2020"
output: pdf_document
---


# Problem 1

**Hypothesis Test** 

$H_o$ : $\mu_A$ = $\mu_B$ = $\mu_C$ = $\mu_D$ = $\mu_E$ = $\mu_F$ \
vs. \
$H_A$ : not $H_o$ 

**Setting up data and ANOVA**

```{r}
obs <- c(73, 75, 69, 68, 69, 80,
         70, 82, 67, 77, 63, 87,
         71, 82, 76, 76, 65, 86,
         65, 80, 74, 73, 65, 84,
         76, 86, 74, 71, 58, 78)
fert <- factor(rep(c('A', 'B', 'C', 'D', 'E', 'F'), times = 5) )


fert_a <- c(73, 70, 71, 65, 76)
fert_b <- c(75,82,82,80,86)
fert_c <- c(69,67,76,74,74)
fert_d <- c(68,77,76,73,71)
fert_e <- c(69,63,65,65,58)
fert_f <- c(80,87,86,84,78)

mean(fert_a)
mean(fert_b)
mean(fert_c)
mean(fert_d)
mean(fert_e)
mean(fert_f)



ybar <- mean(obs)
 

n <- length(obs)


a <- aov(obs~fert)
summary(a)

alpha <- 0.05
qf(1-alpha , df1 = 5, df2 = 24)
```
As seen from the summary of the ANOVA tables above, we can see that the p value is extremely small.

Since P-value is 5.78e-07 and is less than 0.05. We can conclude that there is evidence to indicate that types of fertilizer effect yield at $\alpha$ = 0.05.

$\hat{\mu} = 74$ \
$y_{ij} = \mu + \tau_i + \epsilon_{ij}$ \
$i = 1,2,3,4,5,6$ \
$j = 1,2,3,4,5$ \ 
$$\sum_{i = 1}^6 \tau_i = 0 $$ \
$\epsilon_{11}, \epsilon_{12}, ... ,\epsilon_{65} \stackrel{iid}{\sim} N(0,\sigma^2)$ \


**LSD TEST**


```{r}
library(agricolae)

LSD.test(a,"fert", console = T)

plot(LSD.test(a,"fert"))

```
From the LSD test we can first see that fertilizer F and B are not significantly different. Also fertilizer D,C, and A are not significantly different. 

If we're going for the highest production of wheat, fertilizer F seems like the way to go with an average of 83 bushels of wheat produced with a standard deviation of 3.872983.



**Analyzing Residuals and model adequacy**

```{r}
qqnorm(a$residuals, main = "Normal Q-Q Plot of Residuals")
qqline(a$residuals, col="red")


plot(a$residuals~a$fitted.values,main="Residuals vs. Predicted",xlab="Predicted",ylab="Residuals")


plot(a$residuals~fert,main="Residuals vs. Approach Type",xlab="Approach Type",ylab="Residuals")
```

Looking at the graphs and analyzing the residuals from this experiment, we can see that there is nothing unusual in the residual plots. The normal Q-Q plots of residuals show that the points follow the line.

# Problem 2 

**Hypothesis Test** 

$H_0: (\tau_\beta)_{ij} = 0$ vs. $H_A:(\tau_\beta)_{ij} \ne 0$

**Setting up data and ANOVA**

```{r}
obs <- c(147, 153, 182, 178, 128, 124,
         157, 161, 144, 150, 121, 119,
         132, 126, 156, 162, 109, 107)

temp <- factor(rep(c(15, 70, 125),each = 2, times = 3))

type <- factor(rep(c('I','II', 'III'),each = 6))


twoway.df <- data.frame(obs = obs,  type = type, temp = temp)
xtabs(~ type + temp, data= twoway.df)
twoway.df

mean(obs)

meas.aov <- aov(obs~temp*type)
summary(meas.aov)

qf(1-alpha,4,9)


```
asdfasdfasdfasdfasdfasdfasdfasdfasdfasdfasdfaaaaaaaaaaaaaaaaaaaaaaa

$F_{4,9,0.05} = 3.633089$

# Model
\
\

$Y_{ijk} = \mu + \tau_i + \beta_j + (\tau\beta)_{ij} + \epsilon_{ijk}$ \

$\mu = 142$ \
$\sum_{i=1}^a\tau_i = 0$ \
$\sum_{j=1}^b\beta_j = 0$ \
$\sum_{j=1}^b(\tau\beta)_{ij} = 0$ \
$\epsilon{ijk} \stackrel{iid}{\sim} N(0,\sigma^2)$ \






# Interaction 

$H_0 : (\tau\beta)_{ij} = 0\ for\ all\ i,j \ vs.H_1:at\ least\ one\ (\tau\beta)_{ij} \ne 0$


$Under\ H_0:  \frac {MS_{AB}} {MS_E} \sim F_{(a-1)(b-1),ab(n-1)}$ 

$\frac {300} {11.1} = 27.0 > F_{4,9,0.05} = 3.633089$

Reject null hypothesis because there are significant interaction. 


# Main Effect of Temperature (Factor A)

$H_0 : \tau_i  = 0$ for all i vs.$H_1:not\ H_0$

$Under\ H_0: \frac {MS_A} {MS_E} \sim F_{a-1,ab(n-1)}$


```{r}
qf(0.95,2,4)
```
$\frac {2976} {11.1} = 267.8 >F_{2,4,0.05} = 6.944272$

Reject null hypothesis and conclude that temperature is a significant factor that affects battery life. 




# Main Effect of material type (Factor B)

$H_0 : \beta_j = 0$ fot all $j$  vs.$H_1: not\ H_0$

$Under\ H_0: \frac {MS_B} {MS_E} \sim F_{b-1,ab(n-1)}$


$\frac {600} {11.1} = 54.05405 > F_{2,4,0.05} = 6.944272$

Reject null hypothesis and conclude that material type does affect battery life. 



